---
title: "Data frames"
format:
  live-html:
    theme:
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300
    toc: false
    section-divs: true
    include-after-body:
      - text: |
          <script src="../../js/progressive-reveal.js"></script>
          <script src="../../js/bootstrapify.js"></script>

resources:
  - ../../data/durham_weather.csv

engine: knitr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library(checkdown)
source(here::here("R", "helpers.R"))
```

### Data frames {.no-hide}

Most data you will work with comes as a table: one row per observation, one column per variable. In R, a table is a **data frame**. It works much like a pandas `DataFrame` in Python: under the hood, it is a list of vectors of equal length, one vector per column.

The usual way to get a table into R is from a CSV file, with `read.csv()`:

```r
weather <- read.csv("data/durham_weather.csv")
```

`read.csv()` reads the file and returns a data frame. Nothing prints, because the result is saved as the object `weather`. The path is relative to R's working directory; on this site, the file has already been placed in a `data` folder for you.

The chunk below runs this line for you each time you open the page, so that the later exercises work even if you skip ahead.

```{webr}
#| autorun: true
weather <- read.csv("data/durham_weather.csv")
```

### Load the data

Use `read.csv()` to load `data/durham_weather.csv` into an object called `weather`. Then type `weather` on a new line to print it.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
weather <- read.csv("data/durham_weather.csv")
weather
```

:::

###

That's a lot of output! The data frame has one row for each month from January 1900 to December 2025. Printing the whole of a big table is rarely useful; the next few functions give you a quicker look.

### `head()`

`head()` shows the first six rows of a data frame, and `tail()` the last six. Look at the first rows of `weather`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
head(weather)
```

:::

###

The columns are:

| Column | Meaning |
|--------|---------|
| `year`, `month` | when the record was made |
| `tmax_c` | mean daily maximum temperature that month (°C) |
| `tmin_c` | mean daily minimum temperature that month (°C) |
| `frost_days` | number of days with an air frost |
| `rain_mm` | total rainfall that month (mm) |

### Size and names

`nrow()` counts the rows of a data frame, `ncol()` counts its columns, and `names()` lists the column names. Use them to find out how many months `weather` covers and what its columns are called.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
nrow(weather)
ncol(weather)
names(weather)
```

:::

### `str()`

`str()` (short for "structure") gives a compact summary: the size of the data frame, then one line per column with its type and first few values. Run it on `weather`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
str(weather)
```

:::

###

`int` means integer and `num` means a double: see the [types](../02-programming-basics/05-types.qmd) tutorial. `str()` is the quickest way to check that R read each column as the type you expected. A column of numbers that shows up as `chr` (character) usually means there is some text hiding in it.

### Pick out a column with `$`

`$` extracts one column from a data frame as a vector, just as it extracts an element from a named list:

```r
weather$rain_mm
```

Extract the `tmax_c` column. Then use square brackets to show only its first 12 values: the monthly temperatures for 1900.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
weather$tmax_c
weather$tmax_c[1:12]
```

:::

### Quick check

::: {.callout-note appearance="simple" icon=false .question}

**How many rows does `weather` have?**

```{r echo=FALSE}
check_question(
  answer = "1512",
  options = c("126", "1512", "6"),
  type = "radio",
  button_label = "Submit answer",
  q_id = 1,
  right = "Correct! 126 years × 12 months = 1512 rows. There are 6 columns.",
  wrong = "Not quite. Use nrow(weather)."
)
```
:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("02-summaries.html")
```
