---
title: "Summary statistics"
format:
  live-html:
    theme:
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300
    toc: false
    section-divs: true
    include-after-body:
      - text: |
          <script src="../../js/progressive-reveal.js"></script>
          <script src="../../js/bootstrapify.js"></script>

resources:
  - ../../data/durham_weather.csv

engine: knitr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library("checkdown")
source(here::here("R", "helpers.R"))
```

### Summary statistics {.no-hide}

Once you can pull a column out with `$`, summarising it takes one function call. These are the ones you will use most:

| R | Python (NumPy) | Returns |
|---|----------------|---------|
| `mean(x)` | `np.mean(x)` | the mean |
| `median(x)` | `np.median(x)` | the median |
| `sd(x)` | `np.std(x, ddof=1)` | the sample standard deviation |
| `min(x)`, `max(x)` | `np.min(x)`, `np.max(x)` | the smallest and largest values |
| `summary(x)` | — | minimum, quartiles, median, mean and maximum in one go |

Note that `sd()` uses the *sample* formula (dividing by *n* − 1), whereas NumPy's default divides by *n*.

The chunk below loads the data each time you open this page.

```{webr}
#| autorun: true
weather <- read.csv("data/durham_weather.csv")
```

### Mean and standard deviation

Calculate the mean and standard deviation of `tmax_c`, the monthly maximum temperature.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
mean(weather$tmax_c)
sd(weather$tmax_c)
```

:::

###

The mean is about 12.5 °C and the standard deviation about 5.1 °C. Is that a big spread? It mostly reflects the seasons: winter months are much colder than summer ones.

### Extremes

What were the lowest and highest monthly rainfall totals on record? Use `min()` and `max()`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
min(weather$rain_mm)
max(weather$rain_mm)
```

:::

### `summary()`

`summary()` gives the minimum, quartiles, median, mean and maximum all at once. Try it on `rain_mm`. Then try `summary(weather)`, which summarises every column.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
summary(weather$rain_mm)
summary(weather)
```

:::

### Picking out rows

Often you want a summary of just some of the rows: July only, say. A comparison such as `weather$month == 7` returns a *logical* vector, `TRUE` where the month is July and `FALSE` elsewhere. Putting that vector inside square brackets keeps only the `TRUE` values:

```r
weather$tmax_c[weather$month == 7]
```

Calculate the mean July maximum temperature. Then do the same for January.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
mean(weather$tmax_c[weather$month == 7])
mean(weather$tmax_c[weather$month == 1])
```

:::

### Quick check

::: {.callout-note appearance="simple" icon=false .question}

**On average, how much warmer is a July afternoon in Durham than a January one?** (Round to the nearest degree.)

```{r echo=FALSE}
check_question(
  answer = "About 14 °C",
  options = c("About 5 °C", "About 14 °C", "About 20 °C"),
  type = "radio",
  button_label = "Submit answer",
  q_id = 1,
  right = "Correct! About 19.7 °C in July against 6.1 °C in January.",
  wrong = "Not quite. Subtract the January mean of tmax_c from the July mean."
)
```
:::

### Missing values

This dataset has no gaps, but real data often do. R marks a missing value as `NA`, and any calculation involving `NA` returns `NA`: `mean(c(1, 2, NA))` is `NA`. To skip missing values, add `na.rm = TRUE`: `mean(c(1, 2, NA), na.rm = TRUE)` is `1.5`. Try it.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
mean(c(1, 2, NA))


```

## {{< fa circle-check >}}&ensp;Solution

```r
mean(c(1, 2, NA), na.rm = TRUE)
```

:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("03-plots.html")
```
