---
title: "Plots"
format:
  live-html:
    theme:
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300
    toc: false
    section-divs: true
    include-after-body:
      - text: |
          <script src="../../js/progressive-reveal.js"></script>
          <script src="../../js/bootstrapify.js"></script>

resources:
  - ../../data/durham_weather.csv

engine: knitr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library("checkdown")
source(here::here("R", "helpers.R"))
```

### Base R plots {.no-hide}

Base R can draw quick plots with no packages at all. They are the fastest way to look at your data. (From Week 2 you'll use {ggplot2} for figures you want to show other people.)

| R | Python (Matplotlib) | Draws |
|---|---------------------|-------|
| `plot(x, y)` | `plt.scatter(x, y)` | a scatter plot of `y` against `x` |
| `plot(x, y, type = "l")` | `plt.plot(x, y)` | a line |
| `hist(x)` | `plt.hist(x)` | a histogram |

Arguments set the labels: `xlab` and `ylab` label the axes and `main` sets the title.

The chunk below loads the data each time you open this page.

```{webr}
#| autorun: true
weather <- read.csv("data/durham_weather.csv")
```

### A scatter plot

Plot `tmax_c` (on the *y* axis) against `tmin_c` (on the *x* axis). What relationship do you expect?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
plot(weather$tmin_c, weather$tmax_c)
```

:::

### Labels

Default axis labels such as `weather$tmin_c` are hard to read. Redraw the plot with proper labels, including units, by adding `xlab` and `ylab` arguments. Add a title with `main`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
plot(weather$tmin_c, weather$tmax_c)


```

## {{< fa circle-check >}}&ensp;Solution

```r
plot(weather$tmin_c, weather$tmax_c,
     xlab = "Minimum temperature (°C)",
     ylab = "Maximum temperature (°C)",
     main = "Durham, monthly means, 1900–2025")
```

:::

### A histogram

`hist()` shows how the values of one variable are distributed. Draw a histogram of monthly rainfall, with a proper axis label.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
hist(weather$rain_mm, xlab = "Monthly rainfall (mm)")
```

:::

### Bins

`hist()` chooses the number of bins for you, but the choice changes what you see. The `breaks` argument suggests a different number. Redraw the histogram with `breaks = 5`, then with `breaks = 50`. Which gives the most honest picture?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
hist(weather$rain_mm, xlab = "Monthly rainfall (mm)")


```

## {{< fa circle-check >}}&ensp;Solution

```r
hist(weather$rain_mm, breaks = 5, xlab = "Monthly rainfall (mm)")
hist(weather$rain_mm, breaks = 50, xlab = "Monthly rainfall (mm)")
```

:::

###

Too few bins hide the shape of the distribution; too many turn noise into spikes. There is no single right answer, which is why it is worth trying a few.

### Extension: a warming trend?

*Optional.* Plot `tmax_c` for July only against `year`, using the square-bracket trick from the [summary statistics](02-summaries.qmd) tutorial. Add `type = "l"` to join the points into a line. What do you notice?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
july <- weather$month == 7
plot(weather$year[july], weather$tmax_c[july], type = "l",
     xlab = "Year", ylab = "July maximum temperature (°C)")
```

:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("04-writing-functions.html")
```
