---
title: "Writing functions"
format:
  live-html:
    theme:
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300
    toc: false
    section-divs: true
    include-after-body:
      - text: |
          <script src="../../js/progressive-reveal.js"></script>
          <script src="../../js/bootstrapify.js"></script>

resources:
  - ../../data/durham_weather.csv

engine: knitr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library(checkdown)
source(here::here("R", "helpers.R"))
```

### Writing functions {.no-hide}

So far you have used functions that come with R. You can also write your own. A function bundles up some code, so you can reuse it without retyping it and give it a name that says what it does.

```r
celsius_to_kelvin <- function(temp) {
  temp + 273.15
}
```

Compare this with Python:

```python
def celsius_to_kelvin(temp):
    return temp + 273.15
```

There are three differences:

- In R, a function is an object like any other, so you create one with `function()` and save it with `<-`.
- Curly braces `{ }`, not indentation, mark the body of the function. (Indent anyway: it makes code easier to read.)
- R returns the value of the last line automatically, so you rarely need `return()`.

The chunk below loads the weather data, which a later exercise uses.

```{webr}
#| autorun: true
weather <- read.csv("data/durham_weather.csv")
```

### Write a function

Write the `celsius_to_kelvin()` function above, then use it to convert 15 °C to kelvin.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
celsius_to_kelvin <- function(temp) {
  temp + 273.15
}
celsius_to_kelvin(15)
```

:::

###

Because arithmetic in R works element by element, your function works on a whole vector too: `celsius_to_kelvin(c(0, 15, 100))` returns three values.

### Your turn

Write a function called `mm_to_inches()` that converts a length in millimetres to inches (there are 25.4 mm in an inch). Test it on 25.4, which should return 1.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
mm_to_inches <- function(mm) {
  mm / 25.4
}
mm_to_inches(25.4)
```

:::

### More than one argument

A function can take several arguments, and any of them can have a default value. This function calculates the range of a vector: the difference between its largest and smallest values.

```r
value_range <- function(x, na.rm = FALSE) {
  max(x, na.rm = na.rm) - min(x, na.rm = na.rm)
}
```

Use `value_range()` to find the range of `tmax_c`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
value_range <- function(x, na.rm = FALSE) {
  max(x, na.rm = na.rm) - min(x, na.rm = na.rm)
}


```

## {{< fa circle-check >}}&ensp;Solution

```r
value_range <- function(x, na.rm = FALSE) {
  max(x, na.rm = na.rm) - min(x, na.rm = na.rm)
}
value_range(weather$tmax_c)
```

:::

### Ready for Week 1?

You can now load a CSV file, inspect it, pull out columns, summarise them, plot them and write functions. That is everything the Week 1 exercises assume.

Try the self-test at the end of the [preparation guide](https://geol-skills.github.io/materials/preparation.html#quick-self-test). If you can predict what each line does, you are ready.

##

```{r}
#| echo: false
#| results: asis
create_buttons(NULL)
```
