---
title: "Working with data"
format:
  html:
    toc: false
    section-divs: true
    include-after-body:
      - text: |
          <script src="../../js/progressive-reveal.js"></script>
---

```{r include=FALSE}
source(here::here("R", "helpers.R"))
```

### Real data, base R {.no-hide}

The programming basics tutorial showed you how R works. This tutorial puts that to use on a real dataset: monthly weather records from Durham University Observatory, which the Met Office publishes. You'll learn to:

* **load** a CSV file into a data frame
* **inspect** a data frame, and pull out a column with `$`
* **summarise** a column with `mean()`, `sd()` and `summary()`
* **plot** data with `plot()` and `hist()`
* **write** your own function

These are the skills the Week 1 exercises assume. Everything here uses base R: no packages needed.

::: {.callout-note appearance="simple"}
The data are from the Met Office's [historic station data for Durham](https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/durhamdata.txt), © Crown copyright, reused under the [Open Government Licence v3.0](https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/). See [`data/README.md`](https://github.com/geol-skills/r-primers/blob/main/data/README.md) for how they were prepared.
:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("01-data-frames.html")
```
