---
title: "Objects"
format:
  live-html: 
    theme: 
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="../../js/progressive-reveal.js"></script>

engine: knitr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
source(here::here("R", "helpers.R"))
```

### Objects {.no-hide}

To keep a result for later, save it as an **object** with the assignment arrow `<-`:

```r
x <- sqrt(16)
x
x * 2
```

Read `<-` as "gets": `x` gets the value of `sqrt(16)`. It does the job that `=` does in Python. R also accepts `=`, but `<-` is the convention, and it is what this course uses.

Assigning an object prints nothing. To see what an object holds, type its name. Two things to watch:

- Names are case-sensitive, so `x` and `X` are different objects.
- Assigning to a name that already exists replaces the old value without warning.

Run the chunk below, then add a line that stores `x * 2` in a new object called `y`.

```{webr}
x <- sqrt(16)
x
```

### Object names

You can choose almost any name you like for an object, as long as the name does not begin with a number or a special character like `+`, `-`, `*`, `/`, `^`, `!`, `@`, or `&`. 

For instance, check out this list of some possible object names. Some are okay to use; some are invalid:

- `today`: This is fine
- `1st`: This is **bad**; it starts with a number
- `+1`: This is **bad**; it starts with a special character
- `vars`: This is fine
- `\^_^`: This is **bad**; it starts with a special character
- `foo`: This is fine


### Using objects

In the code chunk below, save the results of `rnorm(100, mean = 100, sd = 15)` to an object named `data`. Then, on a new line, call the `hist()` function on `data` to plot a histogram of the random values.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
data <- rnorm(100, mean = 100, sd = 15)
hist(data)
```

:::

### What if?

What do you think would happen if you assigned `data` to a new object named `copy`, like this? Run the code and then inspect both `data` and `copy`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
data <- rnorm(100, mean = 100, sd = 15)
copy <- data


```

## {{< fa circle-check >}}&ensp;Solution

```r
data <- rnorm(100, mean = 100, sd = 15)
copy <- data
data
copy
```

:::

###

Good job! R saves a copy of the contents in data to copy.

### Datasets

Objects provide an easy way to store datasets in R. In fact, R comes with many toy datasets pre-loaded. Examine the contents of `mtcars` to see a classic toy dataset. Hint: how could you learn more about the `mtcars` object?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
mtcars
```

:::

###

Good job! You can learn more about mtcars by examining its help page with `?mtcars`.

### `rm()`

What if you accidentally overwrite an object? If that object came with R or one of its packages, you can restore the original version of the object by removing your version with `rm()`. Run `rm()` on `mtcars` below to restore the mtcars data set.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
mtcars <- 1
mtcars


```

## {{< fa circle-check >}}&ensp;Solution

```r
mtcars <- 1
mtcars
rm(mtcars)
mtcars
```

:::

###

Good job! Unfortunately, `rm()` cannot help you if you overwrite one of your own objects.

##

```{r}
#| echo: false
#| results: asis
create_buttons("04-vectors.html")
```
