---
title: "Lists (optional)"
format:
  live-html: 
    theme: 
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="../../js/progressive-reveal.js"></script>
          <script src="../../js/bootstrapify.js"></script>

engine: knitr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library(tidyverse)
library(checkdown)

source(here::here("R", "helpers.R"))
```

{{< include ../../_optional.qmd >}}

### Lists {.no-hide}

A **list** is like a vector, but its elements can be anything: a number, some text, a whole vector, even another list. Make one with `list()`, usually naming the elements:

```r
site <- list(name = "Durham", elevation_m = 102, temps = c(3.2, 3.8, 5.9))
```

Take an element out by name with `$` (as in `site$name`) or with double brackets (as in `site[["temps"]]`). Single brackets, as in `site["temps"]`, return a smaller *list* rather than the element itself.

If you have used Python, a named list is close to a `dict`.

Lists matter because a data frame is a list of equal-length vectors, one per column. That is why `df$column` works.

Run the chunk below, then extract the mean of `site$temps`.

```{webr}
site <- list(name = "Durham", elevation_m = 102, temps = c(3.2, 3.8, 5.9))
site$name
site[["temps"]]
```

### Lists vs. vectors

::: {.callout-note appearance="simple" icon=false .question}

**Which data structure(s) could you use to store these pieces of data in the same object? `1001`, `TRUE`, `"stories"`**

```{r lists1, echo=FALSE}
check_question(
  answer = c("a list"), 
  options = c(
   "a vector",
   "a list",
   "neither"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("Correct! Lists can contain elements that are different types.")
)
```

:::


### Make a list

Make a list that contains the elements `1001`, `TRUE`, and `"stories"`. Give each element a name.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
list(number = 1001, logical = TRUE, string = "stories")
```

:::


### Extract an element

Extract the number 1001 from the list below.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
things <- list(number = 1001, logical = TRUE, string = "stories")


```

## {{< fa circle-check >}}&ensp;Solution

```r
things <- list(number = 1001, logical = TRUE, string = "stories")
things$number
```

:::

### Data Frames

You can make a data frame with the `data.frame()` function, which works similar to `c()`, and `list()`. Assemble the vectors below into a data frame with the column names `numbers`, `logicals`, `strings`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
nums <- c(1, 2, 3, 4)
logs <- c(TRUE, TRUE, FALSE, TRUE)
strs <- c("apple", "banana", "carrot", "duck")


```

## {{< fa circle-check >}}&ensp;Solution

```r
nums <- c(1, 2, 3, 4)
logs <- c(TRUE, TRUE, FALSE, TRUE)
strs <- c("apple", "banana", "carrot", "duck")
data.frame(numbers = nums, logicals = logs, strings = strs)
```

:::

###

Good job. When you make a data frame, you must follow one rule: each column vector should be the same length


### Extract a column

Given that a data frame is a type of list (with named elements), how could you extract the strings column of the `df` data frame below? Do it.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
nums <- c(1, 2, 3, 4)
logs <- c(TRUE, TRUE, FALSE, TRUE)
strs <- c("apple", "banana", "carrot", "duck")
df <- data.frame(numbers = nums, logicals = logs, strings = strs)


```

## {{< fa circle-check >}}&ensp;Solution

```r
df$strings
```

:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("07-packages.html")
```
